---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# TidyJaccard

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/tidyjaccard)](https://CRAN.R-project.org/package=tidyjaccard)
<!-- badges: end -->

Effortless all vs all jaccard similarity from tidy dataframes.

## Philosophy

For TidyJaccard to work seamlessly, ensure that your dataframe includes two columns:

1. **sample_identifier**: This column should contain unique identifiers for each sample in your dataset.

2. **feature**: Each row in this column defines an element within the set corresponding to the respective sample identifier. This column plays a pivotal role in calculating Jaccard similarity.

Specific columns don't matter



## Installation

You can install the development version of tidyjaccard like so:

``` r
remotes::install_github('selkamand/tidyjaccard')
```


## Getting Started

Let's walk through a toy example using a dataset describing physical traits of members of two different families:

Family 1: Smith (5 members)
Family 2: Johnson (3 members)

For each member of the family, we've recorded a bunch of physical traits. Each row in our dataset describes one trait present in one individual.

Lets look at the data:

```{r}
# Read data
path_data <- system.file(package='tidyjaccard', 'smith_and_johnson.csv')
df_family <- read.csv(path_data)

# Print data
head(df_family)
```

Now lets compute similarity between each person by their traits

```{r example}
library(tidyjaccard)

df_family |>
  tidy_pairwise_jaccard_similarity()
```
